\section{GraphDrone Fit Phase II-A}

\paragraph{Goal.}
Phase II-A extends the new \texttt{GraphDrone.fit()} package with portfolio-generic geometry experts while keeping the public benchmark surface limited to real model families:
\texttt{GraphDrone}, \texttt{TabPFN}, \texttt{TabR}, and \texttt{TabM}.

\paragraph{Implementation.}
Two internal geometry specialists were added on top of the anchor feature space:
\[
\texttt{GEOMETRY\_1} = \texttt{LID} + \texttt{mean\_knn\_distance},
\qquad
\texttt{GEOMETRY\_2} = \texttt{LOF} + \texttt{mean\_knn\_distance}.
\]
The first implementation reused anchor quality channels for these geometry experts.
That created a routing ambiguity: the router could identify the expert family from descriptor tokens, but not their row-level reliability from the quality channels.

\paragraph{Correction.}
The quality tensor was then extended with geometry-specific channels:
\[
q^{\text{geom}}_{i,e}
=
\big[
\texttt{lid}_{i,e},
\texttt{lof}_{i,e},
\texttt{mean\_knn\_distance}_{i,e}
\big],
\]
and these were appended to the legacy quality features.
The adapter was also hardened against degenerate LID cases so the geometry channels remain finite.

\paragraph{Smoke outcome.}
On the initial registered-portfolio smoke pass, GraphDrone beat TabPFN on six of nine datasets.
Internal diagnostics showed that geometry experts were sometimes individually stronger than the anchor, but the public GraphDrone row often remained at or near the anchor.
After the geometry-quality correction, routing behavior changed materially:
\begin{itemize}
\item specialist identities are now explicit in diagnostics;
\item geometry specialists are now selected in routed solutions;
\item performance improved on \texttt{healthcare\_insurance\_expenses};
\item performance regressed slightly on \texttt{houses} and \texttt{concrete\_compressive\_strength};
\item \texttt{diamonds} stayed flat.
\end{itemize}

\paragraph{Conclusion.}
Phase II-A closes a real implementation gap: geometry experts are no longer present only as detached diagnostics.
The remaining issue is no longer hidden channel aliasing, but routing sensitivity and specialist integration quality.
